Domestic cooking and likelihood

### 2022-10-25 inst/cook.Rmd
more work

```{r loadlib}
library(hyper2)

```

@@

@@

>>>>

```{r loadlib}
library(hyper2)
```

@@

```{r,label=plotsupplog}
plot(log(lam),L-max(L),type="l",lwd=2)
abline(h=c(0,-2),col='gray')
abline(v=1,col='red')
text(0,-0.2,"lambda=1",col="red",pos=4)
```

@@

We can calculate the maximum likelihood estimate as follows:

```{r,label=maxlikelambda,cache=TRUE}
o <- optimize(supplam,c(40,70),maximum=TRUE)
o
```

To calculate a formal likelihood ratio we need to calculate the
support for $\lambda=1$ and compare it with the support for
$\lambda=\hat{\lambda}$

```{r,supmax,cache=TRUE}
supplam(o$maximum)
supplam(1)
```

and we get a $p$-value for $H_0\colon\lambda=1$ using Wilks,
specifically that
$2\log\left(\frac{\mathcal{L}_{\lambda=\hat{\lambda}}}{\mathcal{L}_{\lambda=1}}\right)\sim\chi^2_1$

```{r pvalueforlambdaequalsone,cache=TRUE}
pchisq(2*(supplam(o$maximum)-supplam(1)),lower.tail=FALSE,df=1)
```

Actually we should be able to analyse $\lim_{\lambda\longrightarrow\infty}H_\lambda$:

```{r,lambigcalc,cache=TRUE}
lambig <- 10^(0:6)
LL <- sapply(lambig,supplam)
```

```{r,plotlambdiff,echo=F}
plot(log(lambig),LL-max(LL),pch=16)
for(i in seq(from=0,to=8,by=2)){abline(i,-1,col='gray',lwd=0.5)}
abline(v=log(o$maximum))
```

Looking at the two parts of the likelihood function that includes
lambda, and thinking about $\lim_{\lambda\longrightarrow\infty}$ we
see that the first product is approximately constant and the second is
$\propto\lambda^{-1}$.  This analysis is a little bit naive because as
$\lambda$ grows very large, we might expect $\hat{p_3}$ to grow too.  But:

```{r,maxplam1110,cache=TRUE}
M <- rbind(
maxp(H3(10^0    ),n=1,give=TRUE)$par,
maxp(H3(10^1    ),n=1,give=TRUE)$par,
maxp(H3(10^2    ),n=1,give=TRUE)$par,
maxp(H3(10^3    ),n=1,give=TRUE)$par,
maxp(H3(10^4    ),n=1,give=TRUE)$par,
maxp(H3(10^5    ),n=1,give=TRUE)$par,
maxp(H3(10^6    ),n=1,give=TRUE)$par
)
M
```

```{r}
matplot(M,type='b')
```

Above we see that the evaluate hardly changes when $\lambda > 10^4$,
and indeed from the figure we see that the slope of the support graph
is about $-1$ for large values of $\lambda$ [the points follow a line
parallel to one of the slanted gray lines, which have a slope of
exactly $-1$; note that the aspect ratio of the graph is not unity]
